---
title: "Classical perspectives of controling acquiescence"
description: |
  Explores properties of two methods (centering and standardizing) of recoding scores to control for acquiescence bias.
author:
  - name: Ricardo Primi 
    url: https://www.labape.com.br
    affiliation: USF
    affiliation_url: https://www.usf.edu.br
date: "`r Sys.Date()`"
output: radix::radix_article
---


```{r include=FALSE}

library(radix)

```

  
#### Ten berge partialling the mean (centering method) 

Consider a six-item scale composed by three pairs of antonym itmes scored ina 5-point likert scale. Let $i = 1, 2, 3$ be items phrased on the positive pole, and $i = a,b,c$ be the negative ones and $x_{ij}$ be the original response of subject $j$ on item $i$. The aquiescence index $acq_j$ of a subject $j$ is given by:  
  
  
$$acq_j =\frac{1}{6} \left[\sum_{i=1}^{3}x_{ij} + \sum_{i=a}^{c}x_{ij} \right ]$$  
  
  
The total score $scr_j$ on the scale is given by:  

$$scr_j = \frac{1}{6} \left[\sum_{i=1}^{3}x_{ij} + \sum_{i=a}^{c} \left( 6-x_{jj} \right) \right]$$
  
  
And the recoded score  
$$scr.rec_{j}=\frac{1}{6} \left[  \sum_{i=1}^{3} \left( x_{ij}- acq_j \right) -1 \left(  \sum_{i=a}^{c} \left( x_{ij}-acq_j \right) ~ \right)  \right]$$  
  
  
$$scr.rec_{j}=\frac{1}{6} \left[  \sum_{i=1}^{3}x_{ij} - \sum _{i=1}^{3}acq_j - \sum _{i=a}^{c}x_{ij} + \sum _{i=1}^{3}acq_j \right]$$  
  
  
That simplifies to:  
  
$$scr.rec_{j}=\frac{1}{6} \left[  \sum _{i=1}^{3}x_{ij}- \sum _{i=a}^{c}x_{ij} \right]$$

It is interesting to note that for balanced scales the original scores is a linear transformation of original scores, that is, the classical scores is automatic recoded.  

The score culd be written in this form:  
$$scr_{j}= \frac{1}{6} \left[\sum_{i=1}^{3}x_{ij} + \sum_{i=a}^{c} \left( 6-x_{ij} \right) \right] = 
\frac{1}{6} \sum_{i=1}^{3}x_{ij} + \frac{1}{6} \sum_{i=a}^{c} \left(6-x_{ij} \right) =
\frac{1}{6} \left[  \sum _{i=1}^{3}x_{ij}- \sum _{i=a}^{c}x_{ij} \right] + 3$$ 


That makes explicit that the first component $\frac{1}{6} \left[  \sum _{i=1}^{3}x_{ij}- \sum _{i=a}^{c}x_{ij} \right]$ is equal to $scr.rec_{j}$. Therefore:

$$scr_j =  scr.rec_{j} + 3$$  


#### Simulating balanced scale of 6 items (3 positively keyed and 3 negatively keyed)  


* Produces a dataset 6 items of 5 point likers scale.  
* Create all$5^6=15625$ possible response patterns  



```{r echo=TRUE}
  
  library(crossdes)
  library(tidyverse)

  r_ptrn <- permutations(5, 6, repeats.allowed=TRUE)
  colnames(r_ptrn) <- paste("i", 1:6, sep="")

```
 
* Create mean and standard deviation within subjects and recode responses using two methods: (a) centering $c_{ij} = x_{ij}-acq_j$ and (b) $z_{ij} = (x_{ij}-acq_j)/sd_j$  


```{r echo=TRUE}
 
  acq_j <- apply(r_ptrn, 1, mean)
  sd_j <- apply(r_ptrn, 1, sd)
  
  r_ptrn_c <- apply(r_ptrn, 2, function(x) { x - acq_j} )
  r_ptrn_z <- apply(r_ptrn, 2, function(x) { (x - acq_j)/sd_j} )
  

```

* Calculate scores. Note we reverse three last items as if they were negatively worded items.

```{r echo=TRUE}

  scr_j  <- r_ptrn %>% 
    as.data.frame %>%
    mutate_at(4:6,  funs(6-.)) %>%
    apply(1, mean, na.rm=TRUE)
  
  scr.rec_j  <- r_ptrn_c %>% 
    as.data.frame %>%
    mutate_at(4:6,  funs(-1*.)) %>%
    apply(1, mean, na.rm=TRUE)
  
  scr.z.rec_j  <- r_ptrn_z %>% 
    as.data.frame %>%
    mutate_at(4:6,  funs(-1*.)) %>%
    apply(1, mean, na.rm=TRUE)

```


* Create a tibble with scores acquiescence index  

```{r echo=TRUE}

 df <- tibble(
   scr_j,
   scr.rec_j,
   scr.z.rec_j,
   acq_j,
   sd_j
  )

```


#### Visualizing relationship among scores and acquiescence parameters

* How recoded scores (standardized and centered) are related to original scores
* How sd_j infuences transformation
* What is the relaioship between acq_j (mean) and standard deviation sd_j 



```{r, echo=TRUE}
  
  library(artyfarty)
  library(psych)
  library(knitr)
   
  df %>%
    ggplot(aes(y=scr_j, x=scr_j , colour=acq_j)) + 
    geom_point(alpha = .6) +  
    theme_farty()   +
    scale_colour_gradientn(colours = pal("ashes"))
   
  df %>%
    ggplot(aes(y=scr.z.rec_j, x=scr_j , colour=sd_j)) + 
    geom_point(alpha = .6) +  
    theme_farty()   +
    scale_colour_gradientn(colours = pal("ashes"))
 
   df %>%
    ggplot(aes(y=sd_j, x=acq_j, colour=acq_j)) + 
    geom_point(alpha = .6) +  
    theme_farty()   +
    scale_colour_gradientn(colours = pal("ashes"))

  df %>%
    describe %>%
    kable(digits = 2)

```



