---
title:  "Simulations discussed used in Acquiescence and Person Differential Functioning: Solving Person DIF with Balanced Scales"
author: "Ricardo Primi"
date: "20/07/2018"
output: html_document
theme: paper


---
```{r setup, include=FALSE}
  knitr::opts_chunk$set(echo = TRUE, warnings = FALSE, message = FALSE)
```

#### Libraries
```{r}
  
  library(tidyverse)
  
  library(sjPlot)
  library(artyfarty)
 
  library(mirt)
  library(psych)
  library(knitr)
  library(lavaan)

  library(xlsx)

 
```

#### Loads empirical data

```{r}
 
  bd_bjmsp <-  readRDS("bd_bjmsp.rds")
  item_dic_c <-  readRDS("item_dic_c.rds")
  
 
```
#### Load R functions 

This file contains all functions used in the paper:  
  
* `sim_trait_acq` that simulates two factors trait and acquiescence. It has the following arguments:  
    + N=10000, number of subjects
    + mean_a1, mean of the discrimination distribution
    + sd_a1, standard deviation of the discrimination distribution
    + sd_acq,  ariance of the acquiescence factor
    + n_pw,  number of positively worded items
    + n_nw,  number of negatively worded items
    + interc_pw, matrix with positively worded items itntercepts to draw from
    + interc_nw   matrix with negatively worded items itntercepts to draw from  
    
* `classical_scores` that calculates acquiescence and recoded scores  
* `pdif_scores` that calculates pddif using graded  response model  
* `ri_scores` that calculates multidimensional graded response model with random intercept  

```{r}

 source("simul_functions.R")

```

#### Calculate basic parameters in empirical data

```{r}
 
# recover parameters
  pars3 <- bd_bjmsp %>% 
    select(item_dic_c$item_labels) %>%
    mutate_if(item_dic_c$pole==0, funs(6-.)) %>%
    mirt(2, itemtype = 'graded', pars = 'values',TOL = .001)

# fix loadins 1 and -1 of random intercept factors F2, fix est=F and estimates variance of F2
  pars3[pars3$name=="a2", ]$value <- item_dic_c$a2 
  pars3[pars3$name=="a2", ]$est <- FALSE 
  pars3[pars3$name=="a1", ]$value <- abs(pars3[pars3$name=="a1", ]$value)
  pars3[pars3$name=="COV_22", ]$est <- FALSE 

# calibrate item parameters
  ri <- bd_bjmsp %>% 
    select(item_dic_c$item_labels) %>%
    mutate_if(item_dic_c$pole==0, funs(6-.)) %>%
    mirt(2, pars = pars3,itemtype = 'graded', TOL = .001)

# examines results
  summary(ri)
  coef(ri , simplify=TRUE, irt.parms = TRUE)
  
  coef(ri , simplify=TRUE)$items %>% 
    as.data.frame() %>%
    describe

  coef(ri , simplify=TRUE)$items %>% 
     as.data.frame() %>%
     describeBy( group = .$a2)
  
```


#### Simulate datasets

* 10.000 examinees 
* 12 items scale, 5-likert options
* 3 conditions: Balanced (6 vs 6), Unbalanced Positive (8 vs 4), Unbalanced negative (4 vs 8)
* 50 replications per condition


```{r eval=FALSE}

# colect basic parametrs from real data  

  mean_a1 = mean(coef(ri , simplify=TRUE)$items[, "a1"])
  
  sd_a1 = sd(coef(ri , simplify=TRUE)$items[, "a1"]) 
  
  interc_pw  <-  coef(ri, simplify=TRUE)$items %>% as.data.frame() %>%
          arrange(a2) %>% filter(a2==1) %>% select(3:6)
  
  interc_nw  <-  coef(ri, simplify=TRUE)$items %>% as.data.frame() %>%
          arrange(a2) %>% filter(a2==-1) %>% select(3:6)

# variable simulation parameters
# number of replications
r <- 50  
prms <- 
  tibble(
    n_pw = c(rep(6, r), rep(9, r), rep(3, r)),
    n_nw = c(rep(6, r), rep(3, r), rep(9, r))
  )

# loop to simulate using purrr (cool !!)
 
 dt_simul <- prms %>% pmap(
    sim_trait_acq, 
    N=10000,
    mean_a1 = mean_a1,
    sd_a1 = sd_a1, 
    sd_acq = sqrt(.40),
    interc_pw = interc_pw,   
    interc_nw = interc_nw  
   )


 
```

####  Calculate Acquiescece and Trait from 3 methods: classical, IRT-PDIF and random-intercept

```{r eval=FALSE}

  classical <- map(dt_simul, classical_scores, lik=3)
  pdif <- map(dt_simul, pdif_scores)
  ri_sim <- map(dt_simul, ri_scores)
  ri_sim2 <- map(dt_simul, ri_scores, a="a3")
  
  coef(ri_sim2[[51]][[2]], simplify = TRUE)
  coef(ri_sim2[[101]][[2]], simplify = TRUE)
  
  summary(pdif)
  coef(ri_sim2[[101]][[2]], simplify = TRUE)

```

#### Extract scores and run correlations
```{r eval=FALSE}

# extract data
     pdif_scores <- map(pdif, "scrs")
     ri_scores <- map(ri_sim, "scrs")
     ri_scores2 <- map(ri_sim2, "scrs")
     ri_scores2 <- map(ri_scores2, as.data.frame)
     
     nm <- paste(dimnames(ri_scores2[[2]])[[2]], "b", sep = "")
     ri_scores2 <- lapply(ri_scores2, function(x) setNames(x, nm))

     true_scores <- map(dt_simul, "persons")
    
     scores <- map2(classical, pdif_scores, cbind)
     scores  <- map2(scores, ri_scores, cbind)
     scores <- map2(scores, ri_scores2, cbind)
     scores <- map2(scores, true_scores, cbind)
   
  # calculate correlations
    correlations <- map(scores, cor)
 
 
  # model fit
     fits <- ri_sim %>% map("mirt_obj") %>% map(M2)
     fits0 <- pdif %>% map("mirt_obj") %>% map(M2)
     
     
  # descriptives
  
  descr <- map(scores, describe) %>% 
    map(as.data.frame) %>%
    map(~mutate(.x, vars2=rownames(.x))) %>%
    map(~gather(.x, key = stats, value = value, -vars, -vars2)) %>%
    enframe() %>% 
    unnest() %>%
    mutate(repl = 
    case_when(
          name <=50 ~1,
          name >50 &  name <=100 ~2,
          name >100 ~3)
      ) %>%
    filter(vars2 %in% 
        c("scr", "scr_rec", "acq", 
        "pdif_mirt", "F1", "F2", "F1b", 
        "F2b","f1_1d", "trait_t", "acq_t")
      ) %>%
    filter(stats %in% c("mean", "sd")) %>%
    select(vars2, stats, value, repl) %>%
    group_by(vars2, stats, repl) %>%
    summarise_all(.funs = mean) %>%
    spread(key = stats, value = value) %>%
    arrange(repl)

   
   fit0_descr <- fits0 %>%  enframe() %>% 
    unnest() %>% summarise_all(.funs = mean)
 
  fits_descr <- fits %>% enframe() %>% 
    unnest() %>% summarise_all(.funs = mean) 
   
   
  save(scores, correlations, 
    fits, fits0, descr, 
    fit0_descr, fits_descr, 
    file = "simul_results2.RData")
 
 
```

##### Calculate average of the 50 simulations

```{r eval=FALSE}

# calculate average correlations among matrices 
# https://stackoverflow.com/questions/19218475/element-wise-mean-over-list-of-matrices



    #  Make a 3D array from list of matrices
    
    arr1 <- array( unlist(correlations[1:50]) , c(22,22,50))
    arr2 <- array( unlist(correlations[51:100]) , c(22,22,50))
    arr3 <- array( unlist(correlations[101:150]) , c(22,22,50))
    
    correlations[[1]] - correlations[[6]]
    
    arr1[, , 2]
 
    #  Get mean and sd of third dimension
    tb1m <- apply( arr1 , 1:2 , mean )
    tb2m <- apply( arr2 , 1:2 , mean )
    tb3m <- apply( arr3 , 1:2 , mean )
    
    tb1sd <- apply( arr1 , 1:2 , sd )
    tb2sd <- apply( arr2 , 1:2 , sd )
    tb3sd <- apply( arr3 , 1:2 , sd )
    
  
    dimnames(tb1m) <- dimnames(correlations[[1]])
    dimnames(tb2m) <- dimnames(correlations[[1]])
    dimnames(tb3m) <- dimnames(correlations[[1]])
    
    dimnames(tb1sd) <- dimnames(correlations[[1]])
    dimnames(tb2sd) <- dimnames(correlations[[1]])
    dimnames(tb3sd) <- dimnames(correlations[[1]])
  
  rm(arr1, arr2, arr3, r)
  
  write.xlsx(tb1m, file = "tables2new.xlsx", sheetName = "tb1m", showNA = FALSE, append = TRUE)
  write.xlsx(tb2m, file = "tables2new.xlsx", sheetName = "tb2m", showNA = FALSE, append = TRUE)
  write.xlsx(tb3m, file = "tables2new.xlsx", sheetName = "tb3m", showNA = FALSE, append = TRUE)
  
  write.xlsx(tb1sd, file = "tables2new.xlsx", sheetName = "tb1sd", showNA = FALSE, append = TRUE)
  write.xlsx(tb2sd, file = "tables2new.xlsx", sheetName = "tb2sd", showNA = FALSE, append = TRUE)
  write.xlsx(tb3sd, file = "tables2new.xlsx", sheetName = "tb3sd", showNA = FALSE, append = TRUE)
  
    
  write.xlsx(as.data.frame(descr), file = "tables3new.xlsx", sheetName = "stat") 

```
